---
title: "Effect of Midlife Physical Activity on Cognition in Elderly Ladies"
author: "Josef Mana"
echo: false
warning: false
message: false
date: "`r Sys.Date()`"
title-block-banner: true
format:
  html: 
    self-contained: true
    code-fold: true
    toc: true
    toc-depth: 3
---

```{r}
#| label: envir

source("_targets_packages.R")

```

```{r}
#| label: func

# function for an APA-style table from gt()
apa <- function(x) {
  
  gt( x, groupname_col = c("Variable","y") ) |>
    tab_options(
      
      table.border.top.color = "white",
      heading.title.font.size = px(16),
      column_labels.border.top.width = 3,
      column_labels.border.top.color = "black",
      column_labels.border.bottom.width = 3,
      column_labels.border.bottom.color = "black",
      table_body.border.bottom.color = "black",
      table.border.bottom.color = "white",
      table.width = pct(100),
      table.background.color = "white"
      
    ) |>
    cols_align(align="center") |>
    tab_style(
      
      style = list(
        cell_borders(
          sides = c("top", "bottom"),
          color = "white",
          weight = px(1)
        ),
        cell_text(
          align="center"
        ),
        cell_fill(color = "white", alpha = NULL)
        ),
      
      locations = cells_body(
        columns = everything(),
        rows = everything()
      )

    ) |>
    opt_align_table_header(align = "left") |>
    cols_label(x ~ "Predictor")

}

```

# Assumptions

The analysis consists of a set of GLMs (logistic for binary variables, least squares Gaussian regressions for continuous variables) with adjustment sets derived from causal assumptions depicted in the following DAG:

```{r}
#| label: fig-dag
#| fig-cap: "Directed acyclic graph (DAG) depicting causal assumptions about relations of variables under investigation. The 'S' node represents the **s**election process that is automatically conditioned on in the sapmle (which is being indicated by a square)."

targets::tar_read(dag_plot)
library(targets)
library(gt)

```


# Analysis & Summary of Results

Estimates comprise of a set of marginalised means and their comparisons from GLMs regressing the outcome on the exposure and adjustement set derived from @fig-dag. Interactions of the exposure and all adjustement set variable were included whereas interactions between adjustment sets' variables were not included. The results are presented for midlife-PA as the exposure and several different (types of) outcomes:

```{r}
#| label: tbl-adj
#| tbl-cap: "Adjusted estimates of the effect of midlife physical activity (m-PA) on cognitive, affective, and physical outcomes of interest. Under the hypothesis presented in @fig-dag, these estimates represent unbiased estimates of total causal effect in the current sample."

tar_read(results)$adjusted$mPA |>
  mutate( Variable = if_else(Variable == "cPA", "Physical Activity", Variable) ) |>
  apa()

```

These can be contrasted against estimates of otherwise equivalent GLMs with no covariates (i.e., unadjusted estimates of association between exposure and outcome):

```{r}
#| label: tbl-unadj
#| tbl-cap: "Unadjusted estimates of the effect of midlife physical activity (m-PA) on cognitive, affective, and physical outcomes of interest. Under the hypothesis presented in @fig-dag, these estimates are biased estimates of total causal effect due to confounding by common causes."

tar_read(results)$unadjusted$mPA |>
  mutate( Variable = if_else(Variable == "cPA", "Physical Activity", Variable) ) |>
  apa()
  

```

# Transparency & Reproducibility

The [renv](https://rstudio.github.io/renv/) package was used to create reproducible environment for the project.
To set-up R environment for reproduction of our results, run:

```r
#install.packages("renv")
renv::restore()
```

The [targets](https://docs.ropensci.org/targets/) package was used to create a reproducible analysis pipeline.
The data are not being shared for the time being. If you managed to get them from the main authors, put them into
the `_raw/` folder and run the analyses using the following code:

```r
#install.packages("targets")
targets::tar_make()
```

The current pipeline looks as follows:

```{r}
targets::tar_visnetwork()
```
